Estimating dark matter halo concentration using the integrated mass profile

### 2015-11-07 paper/paper.tex
reescribiendo cada frase (?)

In a dark matter dominated universe galaxy properties are driven by
the dark matter halo host.
One of the key structural halo properties is the concentration. 
[EN QUE MODELOS IMPORTA LA CONCENTRACION?]
For these reasons it is important to have a reliable concentration
estimate from N-body simulations. 
Here we show that the concentration parameter in dark matter halos from
N-body simulations can be reliably estimated from the radially integrated mass
profile.
The main advantage of this approach is that it uses the full
particle information without any binning, which makes it reliable in
cases when low numerical resolution becomes a limitation for other
methods.   
We show the good performance of this method by estimating halo
concentration both on mock and N-body halos. 
We compare these results against two other methods:
maximum radial velocity measurements and radial particle binning to
fit the density profile. 
Tests on  the mock halos show that the accuracy of our method to recover input
concentrations varies with halo resolution. For halos
sampled with $20$ particles our method recovers the input
concentration with $10\%$ accuracy. 
For the maximum radial velocity method it is on the order of $20\%$
while density profile methods perform worst at the low resolution
limit since they require at least 200 particles inside the virial
radius to make a reliable estimate of the density profile. 
For halos with $2\times 10^4$ particles our method achieves an
accuracy of $0.01\%$ while the velocity and density profile methods
achieve $0.1\%$ and $1\%$, respectively. 
We measure the mass-concentration relationship on N-body
data taking care of using halos sampled with at least $200$
particles. We find that at low masses $10^{12}\hMsun <
M<10^{13}\hMsun$ our method yields median concentration values lower
by a $15\%-20\%$ compared to the velocity and density methods ({\bf
  why?}).  
At higher masses $3\times 10^{13}\hMsun < M < 2\times
10^{14}\hMsun$ the three methods give similar results.
\end{abstract}

@@

We use data from the MultiDark cosmological simulation that follows
the non-linear evolution of a dark matter density field sampled with
$2048^3$ particles over a cubic box of $1000$ \hMpc on a side.  The
data is publicly available through \url{http://www.cosmosim.org/}.
More details about the structure of the database and the simulation
can be found in \citep{2013AN....334..691R}.

>>>>

Galaxy properties are driven by its dark matter halo host.
A dark matter halo's concentration is on the key structural
properties. 
[EN QUE MODELOS IMPORTA LA CONCENTRACION?]
This drives the pursue to make the concentration estimates of
simulated halos as reliable as possible. 
Here we show the radially integrated mass profile can be converted
into a powerful tool to estimate concentratios.

The main advantage of this approach is that it uses the full
particle information without any binning, which makes it reliable in
cases when low numerical resolution becomes a limitation for other
methods.   
We show the good performance of this method by estimating halo
concentration both on mock and N-body halos. 
We compare these results against two other methods:
maximum radial velocity measurements and radial particle binning to
fit the density profile. 
Tests on  the mock halos show that the accuracy of our method to recover input
concentrations varies with halo resolution. For halos
sampled with $20$ particles our method recovers the input
concentration with $10\%$ accuracy. 
For the maximum radial velocity method it is on the order of $20\%$
while density profile methods perform worst at the low resolution
limit since they require at least 200 particles inside the virial
radius to make a reliable estimate of the density profile. 
For halos with $2\times 10^4$ particles our method achieves an
accuracy of $0.01\%$ while the velocity and density profile methods
achieve $0.1\%$ and $1\%$, respectively. 
We measure the mass-concentration relationship on N-body
data taking care of using halos sampled with at least $200$
particles. We find that at low masses $10^{12}\hMsun <
M<10^{13}\hMsun$ our method yields median concentration values lower
by a $15\%-20\%$ compared to the velocity and density methods ({\bf
  why?}).  
At higher masses $3\times 10^{13}\hMsun < M < 2\times
10^{14}\hMsun$ the three methods give similar results.
\end{abstract}

@@

We use data from the MultiDark cosmological simulation that follows
the non-linear evolution of a dark matter density field sampled with
$2048^3$ particles over a cubic box of $1000\ \hMpc$ on a side.  The
data is publicly available through \url{http://www.cosmosim.org/}.
More details about the structure of the database and the simulation
can be found in \citep{2013AN....334..691R}.
